---
title: "Obtaining Lyrics Data"
author: "Connor French"
date: "3/21/2021"
output: pdf_document
---

Load packages
```{r, message=FALSE}
library(geniusr)
library(tidyverse)
library(tidytext)
library(textdata)
library(here)
```

Get the lyrics for Buck Meek's album Two Saviors (the "positive" music). NOTE: Follow the [geniusr](https://ewenme.github.io/geniusr/) instructions first to authenticate yourself. Otherwise, you won't be able to access the API.  

```{r}
buck_tracklist <- get_album_tracklist_search("Buck Meek", "Two Saviors")


buck_song_df <- map_df(buck_tracklist$song_lyrics_url, get_lyrics_url) %>%
  group_by(song_name) %>% 
  mutate(line_number = row_number()) %>% 
  ungroup() 
```


Get the lyrics for Full of Hell's album Wheeping Choir (the "negative" music)
```{r}
foh_tracklist <- get_album_tracklist_search("Full of Hell", "Weeping Choir")

foh_song_df <- map_df(foh_tracklist$song_lyrics_url, get_lyrics_url) %>%
  group_by(song_name) %>% 
  mutate(line_number = row_number()) %>% 
  ungroup()
```


Put both into a single df
```{r}
total_df <- bind_rows(buck_song_df, foh_song_df)
```

This is the dataset for the workshop.  
```{r, eval=FALSE}
write_csv(total_df, here("data", "lyrics.csv"))
```


## Example analysis

Get one line per token
```{r}
tidy_df <- total_df %>% 
  unnest_tokens(word, line) 
```


Remove stop words
```{r}
no_stop_df <- tidy_df %>% 
  anti_join(stop_words)
```

Check out the most common words for:

Overall
```{r}
no_stop_df %>% 
  count(word, sort = TRUE)
```

Buck Meek's most common words  

```{r}
no_stop_df %>% 
  filter(artist_name == "Buck Meek") %>% 
  count(word, sort = TRUE)
```


Visualize as a bar chart  

```{r}
no_stop_df %>%
  filter(artist_name == "Buck Meek") %>% 
  count(word, sort = TRUE) %>%
  filter(n > 3) %>% 
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```



Most common words for Full of Hell  

```{r}
no_stop_df %>% 
  filter(artist_name == "Full of Hell") %>% 
  count(word, sort = TRUE)
```

```{r}
no_stop_df %>%
  filter(artist_name == "Full of Hell") %>% 
  count(word, sort = TRUE) %>%
  filter(n > 3) %>% 
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```


### Sentiment analysis

Let's use the AFINN and Bing lexicons to get both a continuous and categorical idea of how these lyrics are structured.  

```{r}
afinn_sent <- get_sentiments("afinn")
bing_sent <- get_sentiments("bing")
```

Now let's join this with the original data frame using an `inner_join`  

```{r}
afinn_df <- no_stop_df %>% 
  inner_join(afinn_sent) %>% 
  mutate(index = row_number()) %>% 
  rename(afinn = value)
```

Visualize positivity of each artist.  

Looks like Full of Hell has relatively more negativity than Buck Meek.  

```{r}
afinn_df %>% 
  group_by(artist_name) %>% 
  mutate(artist_index = row_number(),
         overall_sent = if_else(
           afinn >= 0, "positive", "negative"
         )) %>% 
  ggplot(aes(y = afinn, x = artist_index, fill = overall_sent)) +
  geom_col() +
  facet_wrap(~artist_name, nrow = 2)
```


Aggregating single words may not be good enough to get a sense of the sentiment of a body of text. Each song is divided into sections, like the verse, chorus, etc. (although it is not a perfect divide). Let's group the words by their song and section, then summarize the sentiment of each section by taking the sum.  

```{r, message=FALSE}
afinn_df %>% 
  group_by(artist_name, song_name, section_name) %>% 
  summarize(avg_vibe = sum(afinn)) %>% 
  ggplot(aes(x = avg_vibe, color = artist_name)) +
  geom_density()
```
What about a binary look? We can do this with the Bing lexicon
```{r}
bing_df <- no_stop_df %>% 
  inner_join(bing_sent) %>% 
  mutate(index = row_number())

glimpse(bing_df)
```

```{r}
sent_count_df <- bing_df %>% 
  group_by(artist_name) %>% 
  count(sentiment, sort = TRUE) %>% 
  group_by(artist_name) %>% 
  mutate(prop_sent = round(n / sum(n), 3)) %>% 
  ungroup() 

sent_count_df
```
Barplot to visualize counts.  
```{r}
sent_count_df %>% 
  ggplot(aes(x = artist_name, y = n, fill = sentiment)) +
  geom_col(position = "dodge")
```


Barplot to visualize proportions.  

```{r}
sent_count_df %>% 
  ggplot(aes(x = artist_name, y = prop_sent, fill = sentiment)) +
  geom_col()
```
### More serious data wrangling

This is how text may appear if you don't get to use a fancy API to obtain your data. It's a single block with whitespace characters (`\n`) and extraneous classifiers (`[Verse 1]`, `[Instrumental Break]`). We want to get this into a tidy format, where each word is an observation and we have the line number for each word. To do this, we will use the powerful [stringr](https://stringr.tidyverse.org/) package.   

These are Buck Meek lyrics from the album we analyzed earlier! I scraped these lyrics from the Genius web page, using the [rvest](https://rvest.tidyverse.org/) package. This is effectively what the Genius API does, but the API does some helpful transformation under the hood that we'll do here!  

```{r}
candle_lyrics <- "[Verse 1]\nInnocence is a light beam, you're doing your thing\nWith your arm out your window up Highway 9\nWhen it's too much to handle, burn me a candle\nIf you don't have a candle, let me burn on your mind\n[Verse 2]\nThe song of the sirens caught up with me downwind\nMy nose started bleeding by the second note\nHeaven is a motel with a telephone seashell\nWell, check-out's at eleven, and don't ask for more time\n[Chorus]\nWell, did your eyes change? I remember them blue\nOr were they always hazel?\nStill the same face with a line or two\nThe same love I always knew\n[Verse 3]\nI try not to call, but I think I'm being followed\nIt's been about an hour or so\nI hate for you to hear me scared, otherwise, I'm well\nI guess you're still the first place I go\n[Chorus]\nDid your eyes change? I remember them blue\nOr were they always hazel?\nStill the same face with a line or two\nThe same love I always knew\n[Instrumental Break]\n[Chorus]\nDid your eyes change? I remember them blue\nOr were they always hazel?\nStill the same face with a line or two\nThe same love I always knew"
```


Let's split this into lines. The `\n` whitespace character denotes a line break, so we can delimit the string based on this character!
I'll split this into steps for the workshop
```{r}
candle_lyrics %>%
  str_split("\\n") %>% 
  unlist() %>% 
  str_remove("\\[\\D+\\d+\\]") %>% 
  str_remove("\\[\\D+\\]") %>% 
  na_if("") %>% 
  na.omit() %>% 
  enframe(name = "line_number", value = "line") %>% 
  unnest_tokens(word, line)
```








